\section{Introduction}

\subsection{Context of Underwater Robotics and ROVs}

Underwater robotic systems are widely used to offshore energy, marine science and subsea infrastructure inspection. These platforms fall into two categories: ROVs and AUVs. ROVs are tethered vehicles that remain connected to a surface vessel or platform via a long cable providing power and communications. They are typically teleoperated by a human pilot using live video and sensor feedback, and are well suited to tasks that require real-time human decision-making, complex manipulation or compliance with conservative safety procedures.

In contrast, AUVs are untethered vehicles that is powered by their own battery and rely on onboard autonomy for navigation and decision-making. They are generally used for undewater missions, such as seabed mapping, pipeline inspection or environmental monitoring. The vehicle follows a pre-planned route and returns to the surface to offload data at the end of the mission. These operations often take place in harsh environments and at considerable depth, where divers cannot safely work. The ability to extend the endurance and autonomy of such systems has therefore become an active area of research and development.

\subsection{Why Docking Is Important}

Docking systems address servel challenges of underwater robotic operations. A subsea docking station can provide electrical power for battery recharging and a physical interface for high-bandwidth data transfer. Instead of returning to the surface vessel after each mission, a vehicle can periodically return to the dock, recharge, upload data and receive new tasking. This enables longer missions, reduces turnaround time and allows for persistent presence in the area of interest.

Docking also opens up new modes of operation. A dock can act as a local base station for multiple vehicles, support tool or payload exchange, and serve as a fixed reference point for navigation and calibration. For ROVs in particular, docking mechanisms that allow the vehicle to park safely when not in use can reduce wear on equipment and improve safety in congested subsea work sites.

\subsection{Limitations of Current ROV Operations}

Despite their versatility, current ROV operations face several limitations that motivate this project. Work-class ROVs typically depend on an umbilical for both power and communications. A prime example is recent REV Ocean's Aurora ROV, which weighs 4200kg and able to travel up to 6000km range \cite{REVOcean_2023}. However, the tether restricts operational radius, introduces hydrodynamic drag, and can snag on structures or the seabed. For smaller electric ROVs that rely on onboard batteries, endurance is directly limited by battery capacity and the power required for propulsion, lighting and sensors.

Launch and recovery require a surface vessel, crane or A-frame on the deck and a trained team. Operations are constrained by sea state and weather windows, and each deployment incurs significant mobilisation and vessel time costs. When an ROV must repeatedly return to the surface for charging or data offload, these constraints become even more pronounced.

In the research literature, much of the work on autonomous docking has focused on AUVs docking to static, seabed-mounted stations as they are the most common configuration for current AUV systems. These configurations simplify the problem by assuming a fixed dock pose relative to the environment. In contrast, ROV-related docking is often framed in terms of Tether Management Systems (TMS) or simple garages, and is generally piloted manually. Autonomous docking for ROVs to free-floating or suspended stations has received comparatively little attention.

\subsection{What a Free-Floating Docking Station Is}

This project considers a free-floating docking station concept rather than a fixed seabed installation. In this context, a “free-floating” dock refers to a docking structure that is suspended and mechanically connected to a surface structure via a cable or frame, but is not rigidly attached to a fixed structure. The station may be approximately stationary in a local frame, but it can move in surge, sway, heave and rotation due to waves, currents and motion due to the nature of water currents.

A free-floating station offers several advantages. It can be deployed at a convenient depth for the ROV, relocated with the vessel or between sites, and recovered for maintenance without subsea intervention. It also avoids the need for seabed foundations or long-term moorings, which can be costly and environmentally intrusive. However, the resulting relative motion between the dock and the surrounding water makes guidance and control more challenging than for a static seabed dock.

Designing a docking system for such a platform requires explicit consideration of dock dynamics, tether behaviour and the interaction between ROV and station during the final capture phase.

\subsection{Why Vision-Based Docking Is Difficult}

Vision-based docking relies on cameras to detect visual features on or near the docking station and derive the relative pose between the vehicle and the dock. While cameras are a common tool for visual information due to their low cost, rich information content and suitability for fine alignment, underwater imaging introduces several challenges.

Water turbidity causes light scattering and absorption, which reduce contrast, blur edges and degrades the effective visibility range. The lighting conditions underwater are highly variable: at depth there may be very low ambient light, while near the surface specular reflections/refractions and sun glint can saturate parts of the image.

Additionally, onboard computing resources on small ROVs are limited. Algorithms must operate in real time with bounded latency and power consumption. This is particularly challenging for deep learning-based approaches, which can provide strong detection performance but may require optimisation or model compression for embedded deployment.

\subsection{High-Level Description of the Proposed System}

To address these challenges, this thesis proposes a vision-based docking system that combines turbidity-adaptive image processing, two-level marker-based guidance and motion-compensated control. A front camera on the ROV observes a docking station equipped with both active lights and passive fiducial markers.

At longer range, the system uses LED cues as a robust, high-visibility target for coarse guidance. At close range, it transitions to passive markers such as ArUco or AprilTag high contrast patterns mounted on the docking structure to obtain its precise pose estimates via Perspective-\emph{n}-Point (PnP) algorithms. Image pre-processing and enhancement are tuned to underwater conditions to maximise marker and light visibility without amplifying noise.

Pose estimates are fused over time with inertial measurements to produce a stable relative pose, which is then used by a main controller to execute a sequence of behaviours: search for the dock, approach to the dock, align with the dock axis, and final engagement into the docking structure.

\subsection{Thesis Structure}

This Thesis~A report is organised as follows:
\begin{itemize}
	\item Section~\ref{sec:literature-review} presents a literature review of existing AUV and ROV docking systems, with a particular focus on vision-based techniques, other sensing modalities and known limitations of current approaches.
	\item Section~\ref{sec:research-question} defines the research question and hypotheses in detail, and outlines the proposed methodology, including the perception, estimation and control components.
	\item Section~\ref{sec:project-prep} describes the project-dependent preparations that had been undertaken, including preliminary ArUco-based pose estimation on real images and initial CAD modelling of the docking station concept. This section also summarises planned hardware, budget considerations and key risks.
	\item Finally, Section~\ref{sec:conclusion} concludes the report by summarising the current status, highlighting the main anticipated challenges, and outlining the planned work for Thesis~B and Thesis~C, where the full system will be implemented, tested and evaluated through tank and pool experiments using an ROV platform.
\end{itemize}