% Define the Title of the thesis here, including author and supervisor names
\title{\textbf{Vision-Based Autonomous Docking fro ROVs with Relative Motion Between Dock and Vehicle\\ \textit{Interim Report and Project Plan\vspace{6mm}}}}
\author{Author: Chak Wang Choi\rlap{\textsuperscript{1}} \vspace{2mm} \\ Supervisor: Will Midgley \vspace{4mm} \\ UNSW Sydney, School of Mechanical \& Manufacturing Engineering}
\date{21st November 2025}

% Creates the title defined above
\maketitle

% The abstract can be written here
\begin{abstract}
    Remotely Operated Vehicles (ROVs) are unoccupied, high maneuverability underwater robots operated by a human pilot offboard. They are widely used for subsea inspection, maintenance and data collection tasks across industries. However, working-class ROV operations are constraint by vessel time, weather windows and risky launch/recovery procedures, while battery powered vehicles such as AUVs and hybrid ROVs are limited by their battery endurance due to their energy capacity. To address these challenges, many recent AUVs or industry-class ROVs have explored autonomous docking to an underwater charging/data transfer station as a means to extend mission duration and reduce long-term operational costs. Most prior research, however, focuses on AUVs docking to static station fixed to a ridgid structure, with relatively little work on vision-based docking to free-floating or suspended stations. This thesis investigates the feasibility of an underwater-adaptive vision system that uses a two-level strategy: active light cues for long-range guidance and passive fiducial markers for precise six degrees-of-freedom (DOF) pose estimation. This report presents the problem formulation, preliminary literature review, methodology and project plan. Preliminary work includes a basic ArUco pose-estimation under ROS~2 Humble and initial CAD modelling of a suspended docking station.
\end{abstract}

% The nomenclature section is created here (variables and definitions can be edited below).
\nomenclature[1A]{\(A\)}{amplitude of …}
\nomenclature[1Cp]{\(C_p\)}{…}
\nomenclature[1D]{\(D\)}{…}
\nomenclature[1alpha]{\(\alpha\)}{…}
\nomenclature[2TLA]{\(TLA\)}{…}

\printnomenclature

% The students email is inserted below
\footnotetext[1]{z5413670@ad.unsw.edu.au}
